ODGI: scalable tools for pangenome graphs

### 2021-05-31 manuscript/main.tex
Merge introduction stuff where it belongs
Stacked figure
reworked the Abstract a little, explicitly including the Results

    \abstract{ \textbf{Motivation:} Pangenome graphs provide a
      complete and `lossless' representation of genomic information
      that allows for complex population-sized analysis of genomes
      without introducing reference genome bias. Here we present a
      range of novel tools that makes use of the \odgi\ efficient
      in-memory representation of DNA/RNA graphs. \odgi\ can handle
      thousands of large genomes while maintaining fast execution
      runtime and low memory overhead. \odgi\ includes tools for
      detecting complex regions, extracting loci, exploratory
      analysis, sorting, removing artifacts and visualisation.
      %
      \\
      \textbf{Availability:} \odgi\ is software written in the \Cpp\
      programming language and published under the libre MIT license. Source
      code and user manual can be downloaded from
      \href{https://github.com/pangenome/odgi}{https://github.com/pangenome/odgi}.\\
      %
        \textbf{Contact:}
        \href{andreaguarracino@outlook.com}{andreaguarracino@outlook.com},
        \href{erik.garrison@gmail.com}{erik.garrison@gmail.com}\\ \textbf{Supplementary
          information:} Supplementary data are available at
        \textit{Bioinformatics} online.}

@@

%\enlargethispage{12pt}

    \begin{methods}
        \section{Methods}

@@

%        Text Text Text Text Text Text Text Text Text Text Text Text Text
%        Text Text Text Text Text Text Text Text. Figure~\ref{fig:01}
%        shows that the above method Text Text Text Text Text Text Text
%        Text Text Text Text Text.  \citep{Bag01} wants to know about
%            {\ldots}{\ldots} text follows.
%        \begin{equation}
%            \sum \text{\it x}+ \text{\it y} =\text{\it Z}\label{eq:01}\vspace*{-10pt}
%        \end{equation}
%
%        Text Text Text Text Text Text Text Text Text Text Text Text Text
%        Text Text Text Text Text Text Text Text.
%        Figure~2\vphantom{\ref{fig:02}} shows that the above method Text
%        Text Text Text Text Text Text Text Text Text Text Text.
%        \citealp{Boffelli03} might want to know about text text text text
%        Text Text Text Text Text Text Text Text Text Text Text Text Text
%        Text Text Text Text Text Text Text Text.
%        Figure~2\vphantom{\ref{fig:02}} shows that the above method Text
%        Text Text Text Text Text Text Text Text Text Text Text.
%        \citealp{Boffelli03} might want to know about text text text text
%        Text Text Text Text Text Text Text Text Text Text Text Text Text
%        Text Text Text Text Text Text Text Text.
%        Figure~2\vphantom{\ref{fig:02}} shows that the above method Text
%        Text Text Text Text Text Text Text Text Text Text Text.
%        \citealp{Boffelli03} might want to know about text text text
%        text\vspace*{1pt}
%
%        \begin{itemize}
%            \item for bulleted list, use itemize
%            \item for bulleted list, use itemize
%            \item for bulleted list, use itemize\vspace*{1pt}
%        \end{itemize}
%
%        \begin{enumerate}
%            \item this is item, use enumerate
%            \item this is item, use enumerate
%            \item this is item, use enumerate
%        \end{enumerate}
%
%        \subsection{This is subheading}
%
%        Text Text Text Text Text Text Text Text Text Text Text Text Text
%        Text Text Text Text Text Text Text Text.
%        Figure~2\vphantom{\ref{fig:02}} shows that the above method Text
%        Text Text Text Text Text Text Text Text Text Text Text.
%        \citealp{Boffelli03} might want to know about text text text text
%        Text Text Text Text Text Text Text Text Text Text Text Text Text
%        Text Text Text Text Text Text Text Text.
%
%
%        \enlargethispage{6pt}
%
%        \begin{table}[!t]
%            \processtable{This is table caption\label{Tab:01}} {\begin{tabular}{@{}llll@{}}
%                                                                    \toprule head1 &
%                                                                    head2 & head3 & head4 \\\midrule
%                                                                    row1 & row1 & row1 & row1 \\
%                                                                    row2 & row2 & row2 & row2 \\
%                                                                    row3 & row3 & row3 & row3 \\
%                                                                    row4 & row4 & row4 & row4 \\\botrule
%            \end{tabular}}{This is a footnote}
%        \end{table}

    \end{methods}

@@

    \subsection{Features}

@@

    \begin{figure}[!tpb]%figure1
        \fboxsep=0pt\colorbox{gray}{
            \begin{minipage}[t]{235pt}
                \vbox to 100pt{
                    \vfill\hbox to
                    235pt{\hfill\fontsize{24pt}{24pt}\selectfont FPO\hfill}
                    \vfill
                }
            \end{minipage}}
%\centerline{\includegraphics{fig01.eps}}
        \caption{
            Visualizing the Major Histocompatibility Complex (MHC) \textit{locus} pangenome. A) odgi viz 1D
            visualization of 8 haploid, phased human genome assemblies plus the chm13 and GRCh38 reference genomes.
            Graph nodes’ are arranged from left to right. The colored bars represent the linearized renderings of the
            embedded paths versus the nodes in a binary matrix. The black lines under the paths are the links, which
            represent the graph topology. B) Bandage~\citep{26099265} representation of the graph topology of a
            consensus MHC pangenome graph representing only variations bigger than 100 bps. Gene labels were applied
            exploiting \texttt{odgi position} to liftover gene coordinates from the full graph into the consensus
            graph.
        }\label{fig:01}
    \end{figure}

>>>>

    \abstract{
        \textbf{Motivation:} Pangenome graphs provide a complete representation of
        genomic information that allows for complex population-sized analysis of genomes
        without introducing reference genome bias.
        %
        \\
        \textbf{Results:} Here we present \odgi\, a novel range of tools that makes use of an efficient
        in-memory representation of DNA graphs structured in the variation graph model. \odgi\ can
        handle hundreds of large genomes while maintaining fast execution runtime and low memory overhead.
        \odgi\ includes tools for detecting complex regions, extracting \textit{loci}, exploratory analysis,
        sorting, removing artifacts, and graphs manipulation and visualisation.
        %
        \\
        \textbf{Availability:} \odgi\ is software written in the \Cpp\ programming language and
        published under the libre MIT license. Source code and user manual can be downloaded from
        \href{https://github.com/pangenome/odgi}{https://github.com/pangenome/odgi}.\\
        %
        \textbf{Contact:}
        \href{andreaguarracino@outlook.com}{andreaguarracino@outlook.com},
        \href{erik.garrison@gmail.com}{erik.garrison@gmail.com}\\
        \\textbf{Supplementary information:} Supplementary data are available at \textit{Bioinformatics} online.}

@@

@@

@@

@@

    \begin{figure}
        \begin{minipage}[b]{0.48\textwidth}
            \includegraphics[width=\linewidth]{fig/mhc-align.eps}
            \caption{(a) alignment}
            \label{fig:1}
        \end{minipage}

        \begin{minipage}[b]{0.48\textwidth}
            \includegraphics[width=\linewidth]{fig/mhc-pangenome.eps}
            \caption{(b) Visualizing the Major Histocompatibility Complex
                (MHC) \textit{locus} pangenome. A) odgi viz 1D visualization
                of 8 haploid, phased human genome assemblies plus the chm13
                and GRCh38 reference genomes. Graph nodes’ are arranged
                from left to right. The colored bars represent the
                linearized renderings of the embedded paths versus the nodes
                in a binary matrix. The black lines under the paths are the
                links, which represent the graph topology. B)
                Bandage~\citep{26099265} representation of the graph
                topology of a consensus MHC pangenome graph representing
                only variations bigger than 100 bps. Gene labels were
                applied exploiting \texttt{odgi position} to liftover gene
                coordinates from the full graph into the consensus graph.  }

            \label{fig:2}
        \end{minipage}
    \end{figure}
